Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report

### 2024-03-25 revisions.qmd
add dietary protein model, add depression to immune model, add limitations section and detail to discussion.

#### Hormone control model
There is evidence that testosterone is associated with higher numbers of partners, but mixed evidence that it is related to sexual desire (van Anders, 2013). Likewise, evidence of a relationship between circulating testosterone and handgrip strength is also mixed (Nowak-Kornicka et al., 2020). The hormone control model included serum total testosterone (ng/dL). Since male mean testosterone is an order of magnitude higher than the female mean, testosterone was highly confounded with sex. We therefore computed a sex-specific testosterone value as follows: for each sex we centered testosterone at the sex-specific mean and divided by two times the sex-specific standard deviation. This differed from other centered and standardized variables which we centered and standardized across all male and female values.

@@

getncon <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(age_first_sex,
                sex_partners,
                sex_partners_year,
                strength,
                age,
                bmi,
                height,
                weight,
                whitebloodcell,
                hemoglobin,
                testosterone, 
                chronic_disease_score,
                physical_disease_count,
                depression,
                avgcalories, 
                foodinsecurity_adult, 
                tot_MET),
    statistic = list(
      all_continuous() ~ "{N_nonmiss_unweighted}",
      all_categorical() ~ "{N_nonmiss_unweighted}"
    ),
    digits = all_continuous() ~ 0,
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous", 
      foodinsecurity_adult ~ "continuous"
    ),
    missing = "no"
  )

getratio <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(age_first_sex,
                sex_partners,
                sex_partners_year,
                strength,
                age,
                bmi,
                height,
                weight,
                whitebloodcell,
                hemoglobin,
                testosterone, 
                chronic_disease_score,
                physical_disease_count,
                depression,
                avgcalories, 
                foodinsecurity_adult, 
                tot_MET),
    statistic = list(
      all_continuous() ~ "{mean}",
      all_categorical() ~ "{p}"
    ),
    digits = list(
      all_continuous() ~ c(2)
                  ), 
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous",
      foodinsecurity_adult ~ "continuous"
    ),
    missing = "no"
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
  )

con <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(
      age_first_sex,
      sex_partners,
      sex_partners_year,
      strength,
      age,
      bmi,
      height,
      weight,
      whitebloodcell,
      hemoglobin,
      testosterone, 
      chronic_disease_score,
      physical_disease_count,
      depression,
      avgcalories, 
      foodinsecurity_adult, 
      tot_MET
    ),
    label  = list(
      age_first_sex = "Age at first sex (years)",
      sex_partners = "Lifetime number of sexual partners",
      sex_partners_year = "Past year number of sexual partners",
      strength = "Combined Grip Strength (kg)",
      age = "Age (Years)",
      bmi = "Body mass index (kg/m^2)",
      height = "Height (cm)",
      weight = "Weight (kg)",
      whitebloodcell = "White blood cell count (1000 cells/µL)",
      hemoglobin = "Hemoglobin (g/dL)",
      testosterone = "Testosterone (ng/dL)",
      chronic_disease_score = "Chronic Disease Score (0-6)",
      physical_disease_count = "Disease Impairment Score (0-5)",
      depression = "Depression Score (0-27)",
      avgcalories = "Dietary energy intake (kcals)", 
      foodinsecurity_adult = "Food Insecurity Rating (1-4)", 
      tot_MET = "Total MET"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n_unweighted} ({p}%)"),
    digits = all_continuous() ~ 2,
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous", 
      foodinsecurity_adult ~ "continuous"
    ),
    missing = "no"
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
      dplyr::mutate(n_male = getncon$table_body$stat_1, digits = 2) %>%
      dplyr::mutate(n_female = getncon$table_body$stat_2, digits = 2) %>%
      dplyr::relocate(n_male, .before = stat_1) %>%
      dplyr::relocate(n_female, .before = stat_2)
  )  %>%
  add_difference(test = everything() ~ "smd",
                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
  modify_column_hide(columns = ci) %>%
  modify_header(
    label = "**Variable**",
    ratio = "**Ratio**",
    n_male = "**N**",
    n_female = "**N**",
    stat_1 = "**Mean (SD)**",
    stat_2 = "**Mean (SD)**",
    estimate = "*d*"
  ) %>%
  modify_spanning_header(list(
    c("stat_1", "n_male") ~ "**Male**",
    c("stat_2", "n_female") ~ "**Female**",
    c("ratio", "estimate") ~ "**Sexual Dimorphism**"
  )) %>%
  modify_footnote(all_stat_cols() ~ "Weighted means and standard deviations shown for continuous variables, unweighted n (%) for categorical variables") %>%
  modify_caption("**Sex differences for participants ages 18-60 using population weights**") %>%
  bold_labels()

@@

#### 2.2.4 Model of energy intake

Finally, we tested whether strength was related to increased energy intake (Figure \@ref(fig:dietary)). We found that strength was a significant, positive predictor of energy intake controlling for sex. In the expanded control model we found the same pattern while controlling for additional variables.

@@

@@

@@

```{r, vaginalsex}

@@

>>>>

#### Hormone control model
There is evidence that testosterone is positively associated higher numbers of partners men and less robustly so for women (Pollet et al., 2011; van Anders et al., 2007), but mixed evidence that it is related to sexual desire (van Anders, 2013). For review of mixed evidence of the often complex relationships between testosterone and sexual desire, sexual behavior, partnered status, and parenting for men and women see van Anders, 2013. Likewise, although some studies have found a positive correlation between circulating testosterone and handgrip strength (Chiu et al., 2020; Gurlek et al., 2018) other studies have found no association (Gettler et al., 2010; Ribeiro et al., 2016). There is evidence that testosterone-related single nucleotide polymorphisms (SNPs) are associated with greater strength performance (Guilherme et al., 2022). The hormone control model included serum total testosterone (ng/dL). Since male mean testosterone is an order of magnitude higher than the female mean, testosterone was highly confounded with sex. We therefore computed a sex-specific testosterone value as follows: for each sex we centered testosterone at the sex-specific mean and divided by two times the sex-specific standard deviation. This differed from other centered and standardized variables which we centered and standardized across all male and female values.

@@

getncon <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(age_first_sex,
                sex_partners,
                sex_partners_year,
                strength,
                age,
                bmi,
                height,
                weight,
                whitebloodcell,
                hemoglobin,
                testosterone, 
                chronic_disease_score,
                physical_disease_count,
                depression,
                avgcalories, 
                avgprotein, 
                foodinsecurity_adult, 
                tot_MET),
    statistic = list(
      all_continuous() ~ "{N_nonmiss_unweighted}",
      all_categorical() ~ "{N_nonmiss_unweighted}"
    ),
    digits = all_continuous() ~ 0,
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous", 
      foodinsecurity_adult ~ "continuous"
    ),
    missing = "no"
  )

getratio <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(age_first_sex,
                sex_partners,
                sex_partners_year,
                strength,
                age,
                bmi,
                height,
                weight,
                whitebloodcell,
                hemoglobin,
                testosterone, 
                chronic_disease_score,
                physical_disease_count,
                depression,
                avgcalories, 
                avgprotein,
                foodinsecurity_adult, 
                tot_MET),
    statistic = list(
      all_continuous() ~ "{mean}",
      all_categorical() ~ "{p}"
    ),
    digits = list(
      all_continuous() ~ c(2)
                  ), 
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous",
      foodinsecurity_adult ~ "continuous"
    ),
    missing = "no"
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(ratio = round(as.numeric(stat_1)/as.numeric(stat_2), digits = 2))
  )

con <- designsG$d.design.adults %>%
  tbl_svysummary(
    by = sex,
    include = c(
      age_first_sex,
      sex_partners,
      sex_partners_year,
      strength,
      age,
      bmi,
      height,
      weight,
      whitebloodcell,
      hemoglobin,
      testosterone, 
      chronic_disease_score,
      physical_disease_count,
      depression,
      avgcalories, 
      avgprotein,
      foodinsecurity_adult, 
      tot_MET
    ),
    label  = list(
      age_first_sex = "Age at first sex (years)",
      sex_partners = "Lifetime number of sexual partners",
      sex_partners_year = "Past year number of sexual partners",
      strength = "Combined Grip Strength (kg)",
      age = "Age (Years)",
      bmi = "Body mass index (kg/m^2)",
      height = "Height (cm)",
      weight = "Weight (kg)",
      whitebloodcell = "White blood cell count (1000 cells/µL)",
      hemoglobin = "Hemoglobin (g/dL)",
      testosterone = "Testosterone (ng/dL)",
      chronic_disease_score = "Chronic Disease Score (0-6)",
      physical_disease_count = "Disease Impairment Score (0-5)",
      depression = "Depression Score (0-27)",
      avgcalories = "Dietary energy intake (kcals)", 
      avgprotein = "Dietary protein intake (grams)",
      foodinsecurity_adult = "Food Insecurity Rating (1-4)", 
      tot_MET = "Total MET"
    ),
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n_unweighted} ({p}%)"),
    digits = all_continuous() ~ 2,
    type = list(
      chronic_disease_score ~ "continuous",
      physical_disease_count ~ "continuous", 
      foodinsecurity_adult ~ "continuous"
    ),
    missing = "no"
  ) %>%
  modify_table_body(
    ~ .x %>%
      dplyr::mutate(ratio = getratio$table_body$ratio, digits = 2) %>%
      dplyr::mutate(n_male = getncon$table_body$stat_1, digits = 2) %>%
      dplyr::mutate(n_female = getncon$table_body$stat_2, digits = 2) %>%
      dplyr::relocate(n_male, .before = stat_1) %>%
      dplyr::relocate(n_female, .before = stat_2)
  )  %>%
  add_difference(test = everything() ~ "smd",
                 estimate_fun = purrr::partial(style_ratio, digits = 2)) %>%
  modify_column_hide(columns = ci) %>%
  modify_header(
    label = "**Variable**",
    ratio = "**Ratio**",
    n_male = "**N**",
    n_female = "**N**",
    stat_1 = "**Mean (SD)**",
    stat_2 = "**Mean (SD)**",
    estimate = "*d*"
  ) %>%
  modify_spanning_header(list(
    c("stat_1", "n_male") ~ "**Male**",
    c("stat_2", "n_female") ~ "**Female**",
    c("ratio", "estimate") ~ "**Sexual Dimorphism**"
  )) %>%
  modify_footnote(all_stat_cols() ~ "Weighted means and standard deviations shown for continuous variables, unweighted n (%) for categorical variables") %>%
  modify_caption("**Sex differences for participants ages 18-60 using population weights**") %>%
  bold_labels()

@@

#### 2.2.4 Models of energy and protein intake

Finally, we tested whether strength was related to increased energy (Figure \@ref(fig:dietary)) or protein intake (Figure \@ref(fig:protein)). We found that strength was a significant, positive predictor of both energy and protein intake controlling for sex. In the expanded control model we found the same pattern while controlling for additional variables.

@@

```{r protein}
Figure5 <- forestplot(
  m_protein,
  m_protein_alt,
  intercept = F,
  facet = F,
  dodgewidth = .4,
  varnames = vnames, 
  modelnames = mmnames) +
  theme_minimal(17) +
  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T))

ggsave("Figures/Figure5.pdf", Figure4, width = 14, height = 7)

Figure5 
```

@@

We now argue that these results emphasize the importance of biparental care in human reproduction (Gettler et al., 2020; Bribiescas et al., 2012). Across evolutionary time, fitness for both men and women was heavily dependent on obtaining successful long term partners, thereby reaping the efficiency benefits of a sexual division of labor (Quinlan, 2008; Quinlan & Quinlan, 2007). More specifically, humans require provisioning assistance for an extended period (Insert figure?) (Davison & Gurven, 2022; Davison & Gurven, 2021) and paternal care is an important predictor of child outcomes (Starkweather et al., 2021; Winking et al., 2011). Thus, our pilot study refines the sexual selection hypothesis for the relationship between strength and mating success to emphasize that mating success is best operationalized as long-term mating, rather than numbers of sexual partners. Moreover, because strength indexes productivity in addition to formidability, our results could demonstrate female choice for productivity and therefore the ability of stronger men to provision offspring through big-game hunting (Kaplan et al., 2000).

### Limitations
Our registered report has one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is greater for males than females. Our aim was not to test all possible byproduct hypotheses, nor is it possible to do so. Therefore, we are unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS). Likewise, NHANES does not collect data on sociosexuality. Likewise, as described in our measures, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from sexual behavior for both men and women. Measures of mating success, including number of sexual partners and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions (Pérusse, 1993).

@@

```{r vaginalsex}

@@

```{r corrmatf}
corr_f <- corrplot(cor_mat_cors_f, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
                   title = "Correlation matrix of coninuous study variables, Females 18-60")

```{r corrmatm}

corr_m <- corrplot(cor_mat_cors_m, method = 'shade', addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75,
                   title = "Correlation matrix of coninuous study variables, Males 18-60")

```{r corrmatcom}

corr_com <- corrplot(cor_mat_cors, addCoef.col = "black", number.digits = 2, type = "lower", number.cex = 0.75, 
title = "Correlation matrix of coninuous study variables, Males and Females 18-60")
```
